Named Entity Recognition as Structured Prediction

### 2012-11-08 doc/report/paper.tex
Some changes, abstract placeholder.

% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition with Structured Prediction}

@@

\author{Joachim Daiber \\
  University of Groningen \\
  %Affiliation / Address line 2 \\
  %Affiliation / Address line 3 \\
  {\tt jodaiber@gmx.de} \\\And
  Carmen Klaussner \\
  University of Groningen \\
  %Affiliation / Address line 2 \\
  %Affiliation / Address line 3 \\
  {\tt Carmen@wordsmith.de} \\}

@@

\section{Introduction}
Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
Table \ref{table:NETypes} gives an overview of the different NE types.

@@

\subsection*{Issues in Named Entity Recognition}
Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{nadeau2007survey}, 
the property of some lexical representation having
more than one possible meaning, and \emph{Metonymy}, which refers to the concept of the part-whole/whole-part relation between two expressions. 
\emph{Polysemy} could become an issue for NER when the lexical representation of an item could point to two different \emph{NE} types.
This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
Often the context will not be disambiguating as in \ref{ex:Paris}.

@@

\subsection*{Methods employed for NER}
For \emph{NER}, there exist both rule-based and statisical approaches. 
Rule-based methods make use of the underlying rules governing languages to extract named entities. 
However, this approach is high maintenance, quite time-consuming and requires extensive work of computational linguists \cite{nadeau2007survey}
and although the results are often high in precision, lacks considerably in recall.

@@

\subsection*{CoNLL Data and Baseline Performance}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.

>>>>

% \usepackage{xcolor}
% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
% \definecolor{medium-blue}{rgb}{0,0,0.5}
% \hypersetup{
%     colorlinks, linkcolor={dark-red},
%     citecolor={dark-blue}, urlcolor={medium-blue}}
\title{Named Entity Recognition as Structured Prediction}

@@

\author{Joachim Daiber \\
  University of Groningen \\
  2397331\\
  %Affiliation / Address line 3 \\
  {\tt jodaiber@gmx.de} \\\And
  Carmen Klaussner \\
  University of Groningen \\
  2401541\\
  %Affiliation / Address line 3 \\
  {\tt Carmen@wordsmith.de} \\}

@@

\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
\end{abstract}

\section{Application: Named Entity Recognition}

Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
Table \ref{table:NETypes} gives an overview of the different NE types.

@@

\subsection{Issues in Named Entity Recognition}
Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{nadeau2007survey}, 
the property of some lexical representation having
more than one possible meaning, and \emph{Metonymy}, which refers to the concept of the part-whole/whole-part relation between two expressions. 
\emph{Polysemy} could become an issue for NER when the lexical representation of an item could point to two different \emph{NE} types.
This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
Often the context will not be disambiguating as in \ref{ex:Paris}.

@@

\subsection{Methods employed for NER}
For \emph{NER}, there exist both rule-based and statisical approaches. 
Rule-based methods make use of the underlying rules governing languages to extract named entities. 
However, this approach is high maintenance, quite time-consuming and requires extensive work of computational linguists \cite{nadeau2007survey}
and although the results are often high in precision, lacks considerably in recall.

@@

\subsection{CoNLL Data and Baseline Performance}
The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
% A baseline rate was computed for the English and the
% German test sets. It was produced by a system which
% only identified entities which had a unique class in
% the training data. If a phrase was part of more than
% one entity, the system would select the longest one.
% All systems that participated in the shared task have
% outperformed the baseline system.
% For all the Fβ=1 rates we have estimated sig-
% nificance boundaries by using bootstrap resampling
% (Noreen, 1989). From each output file of a system,
% 250 random samples of sentences have been chosen
% and the distribution of the Fβ=1 rates in these sam-
% ples is assumed to be the distribution of the perfor-
% mance of the system. We assume that performance
% A is significantly different from performance B if A
% is not within the center 90% of the distribution of B.
